---
title: "Template RST South Sudan"
output: 
 html_document:
    code_folding: "hide"
date: "`r Sys.Date()`"
params:
  file_path:
    label: "Please provide the path to the file with your linelist data"
    value: "data/mpox_data_gh_burundi.csv"
  disease_name: 
    label: "Name of the disease of interest"
    value: "Mpox"
  date_column:
    label: "Please indicate the name of the column in your dataset that contains the dates when cases were identified"
    value: "Date_confirmation"
  deaths_column:
    label: "Please indicate the name of the column in your dataset that contains the dates when deaths occurred"
    value: "Death_date"
  reporting_delay:
    label: "Please provide the estimated number of days between disease onset and case reporting (if unknown, leave as NA)"
    value: "2"
    choices: ["NA", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10"]
---


```{r settings, echo = FALSE}
knitr::opts_chunk$set(
  fig.width = 9,
  fig.height = 5,
  dpi = 90,
  collapse = TRUE,
  message = FALSE,
  warning = FALSE,
  out.width = "100%"
)
```

```{r, echo=FALSE, include=FALSE}
renv::restore()
```
## Loading packages

``` {r}
library(dplyr)
library(incidence2)
library(epiparameter)
library(EpiNow2)
library(cfr)
library(ggplot2)
library(knitr)
```

## Reading in data

```{r}
# Load in data
linelist <- read.csv(params$file_path)
```


## Converting linelist to incidence

```{r}
incidence_data <- linelist %>%
  incidence2::incidence(date_index = params$date_column,
                        interval = "day",
                        date_names_to = "date",
                        count_values_to = "confirm") %>%
  select(-count_variable)

ggplot(incidence_data, aes(x = date, y = confirm)) +
  geom_col(fill = "darkblue") +
  labs(x = NULL, y = "Case incidence") +
  scale_x_date(date_breaks = "10 days", date_labels = "%b %d") +
  theme_minimal() 

```

# Serial interval

```{r}
si_epidist <- epiparameter_db(
  disease = params$disease_name,
  epi_dist = "serial interval",
  single_epiparameter = TRUE
)

# Discretising and formatting for EpiNow2
si_epidist_disc <- epiparameter::discretise(si_epidist)
si_params <- epiparameter::get_parameters(si_epidist_disc)
si_max <- quantile(si_epidist_disc, 0.999)

si_for_epinow <- Gamma(
  shape = si_params[[1]],
  scale = si_params[[2]],
  max = si_max
)

plot(si_epidist_disc)
```
The mean serial interval period for `r params$disease_name` is of `r si_epidist$summary_stats$mean` days, as described in the literature by `r  si_epidist$citation$author[1]`  et al `r si_epidist$citation$year`.

# Incubation period

```{r}
ip_epidist <- epiparameter_db(
  disease = params$disease_name,
  epi_dist = "incubation",
  single_epiparameter = TRUE
)

# Discretising and formatting for EpiNow2
ip_disc <- epiparameter::discretise(ip_epidist)
ip_params <- epiparameter::get_parameters(ip_epidist)
ip_max <- quantile(ip_epidist, 0.999)

ip_for_epinow <- LogNormal(
  meanlog = ip_params[[1]],
  sdlog = ip_params[[2]],
  max = ip_max
)

plot(ip_disc)
```
The mean incubation period for `r params$disease_name` is of `r ip_epidist$summary_stats$mean` days, as described in the literature by `r  ip_epidist$citation$author[1]`  et al `r ip_epidist$citation$year`.

# Estimating infections

```{r, eval=params$reporting_delay == "NA", include=params$reporting_delay == "NA", warning=FALSE, message=FALSE}
epinow_estimates <- epinow(
  data = incidence_data, # time series data
  generation_time = generation_time_opts(si_for_epinow), # assume generation time = serial interval
  delays = delay_opts(ip_for_epinow), # delay from infection-to-onset
  obs = obs_opts(na = "accumulate"),
  rt = NULL
)
epinow_estimates$plots$infections
```


```{r, eval=params$reporting_delay != "NA", include=params$reporting_delay != "NA"}

rep_delay_params <- epiparameter::convert_summary_stats_to_params(
  "lnorm", mean = as.numeric(params$reporting_delay), sd = 0.5
)
rep_delay_for_epinow <- LogNormal(
  meanlog = rep_delay_params$meanlog,
  sdlog = rep_delay_params$sdlog,
  max = 20 
)
epinow_estimates <- epinow(
  data = incidence_data, # time series data
  generation_time = generation_time_opts(si_for_epinow), # assume generation time = serial interval
  delays = delay_opts(ip_for_epinow + rep_delay_for_epinow), # delay from infection-to-onset
  obs = obs_opts(na = "accumulate"),
  rt = NULL
)
epinow_estimates$plots$infections
```

```{r}
epinow_estimates$summary %>% 
  select(measure, estimate) %>%
  kable(caption = "Summary of Estimates")
```




